Notes on QM in generalized coordinates

### 2023-01-21 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
More details on assumptions used

\section{Change of coordinates in manifolds}

\begin{equation}
	\begin{aligned}
		\hat{x}^i &= \hat{x}^i(x^j) \\
		d\hat{x}^i &= \partial_j \hat{x}^i dx^j \\
		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n
	\end{aligned}
\end{equation}

@@

\section{Change of coordinates for wave-function}
\begin{equation}
	\begin{aligned}
		\rho(x^j) &= \psi^\dagger(x^j) \psi(x^j)\\
		||J|| \hat{\rho}(\hat{x}^i) &= || \partial_j \hat{x}^i || \hat{\psi}^\dagger(\hat{x}^i) \hat{\psi}(\hat{x}^i) = \psi^\dagger(x^j) \psi(x^j) = \rho(x^j) \\
		\sqrt{|| \partial_j \hat{x}^i ||} \hat{\psi}(\hat{x}^i) &= \psi(x^j) \\
		\hat{\psi}(\hat{x}^i) &= \sqrt{|| \partial_i x^j ||} \psi(x^j) \\
	\end{aligned}
\end{equation}
Note that $\rho$ can't be negative. Mathematically,  $\rho$ should be the component of a density. Therefore $||J||$ represents the absolute value of the Jacobian determinant.

\section{Change of coordinates in symplectic manifolds}
\begin{equation}
	\begin{aligned}
		\hat{p}_i &= \partial_i x^j p_j \\
		d \hat{p}_i &= \partial_i x^j d p_j \\
		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
	\end{aligned}
\end{equation}

@@

\section{Change of Fourier transform}
\begin{equation}
	\begin{aligned}
		F(\hat{\psi}(\hat{x}^i)) &= F\left(\sqrt{|| \partial_i x^j ||} \psi(x^j)\right) = F\left(\sqrt{|| \partial_i x^j ||}\right) * F \left( \psi(x^j)\right)  \\
	\end{aligned}
\end{equation}
TODO: think of cross-checks

>>>>

\section{Change of coordinates in manifolds}
Transformation rules for differentials on a generic manifold. Only differentiable structure is required.
\begin{equation}
	\begin{aligned}
		\hat{x}^i &= \hat{x}^i(x^j) \\
		d\hat{x}^i &= \partial_j \hat{x}^i dx^j \\
		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
	\end{aligned}
\end{equation}

Transformation rules for a density $\rho(x^i)$. The requirement is that the integral is invariant.
\begin{equation}
	\begin{aligned}
		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n \\
		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
	\end{aligned}
\end{equation}

@@

\section{Change of coordinates for wave-function}

Deducing transformation rules for wave-function based on the fact that the modulues of a wave-function is a density over the manifold.
\begin{equation}
	\begin{aligned}
		\rho(x^j) &= \psi^\dagger(x^j) \psi(x^j)\\
		||J|| \hat{\rho}(\hat{x}^i) &= || \partial_j \hat{x}^i || \hat{\psi}^\dagger(\hat{x}^i) \hat{\psi}(\hat{x}^i) = \psi^\dagger(x^j) \psi(x^j) = \rho(x^j) \\
		\sqrt{|| \partial_j \hat{x}^i ||} \hat{\psi}(\hat{x}^i) &= \psi(x^j) \\
		\hat{\psi}(\hat{x}^i) &= \sqrt{|| \partial_i x^j ||} \psi(x^j) \\
	\end{aligned}
\end{equation}
Note that $\rho$ can't be negative. Mathematically,  $\rho$ should be the component of a density. Therefore $||J||$ represents the absolute value of the Jacobian determinant.

\section{Change of coordinates for momentum}

These are the change of coordinates for momentum. They are derived by making an arbitrary change on the position coordinates and inducing a coordinate change on momentum such that the symplectic form $\omega = dx^i \wedge dp_i$ is invariant. Therefore this depends on the symplectic structure. This makes conjugate momentum vary as a covector.

\begin{equation}
	\begin{aligned}
		\hat{p}_i &= \partial_i x^j p_j \\
		d \hat{p}_i &= \partial_i x^j d p_j \\
		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
	\end{aligned}
\end{equation}

Since in quantum mechanics we do not have a symplectic structure, we can find the same transformation rule based on the fact that $P_i = -\imath \hbar \partial_i$
\begin{equation}
	\begin{aligned}
		\hat{P}_i &= -\imath \hbar \partial_i = -\imath \hbar \partial_i x^j \partial_j = \partial_i x^j P_j \\
	\end{aligned}
\end{equation}

@@

\section{Change of Fourier transform}
\begin{equation}
	\begin{aligned}
		F(\hat{\psi}(\hat{x}^i)) &= F\left(\sqrt{|| \partial_i x^j ||} \psi(x^j)\right) = F\left(\sqrt{|| \partial_i x^j ||}\right) * F \left( \psi(x^j)\right)  \\
	\end{aligned}
\end{equation}
This expression depends only on the definition and properties of the Fourier transform and the transformation rules of densities of manifolds.

TODO: Need a cross check with the transformation rule of momentum
